---
title: "Week 8 Quiz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Inference for one mean

**Finding t using R**

Finding t* using the qt() function:

a. A confidence level of 90%, and df = 10

```{r}
  sample1 <- qt(p=0.10/2, df=10, lower.tail=FALSE)
  print(round(sample1, 2))
```
b. A confidence level of 98% and n = 15

```{r}
  sample2 <- qt(p=0.02/2, df=14, lower.tail=FALSE)
  print(round(sample2, 2))
```

***

>**Confidence interval example**
>
>  Consumer Reports tested 14 brands of vanilla yoghurt and found the following calories per 200g serving:
>  
>  160  200  220  230  120  180  140  130  170  190  80  120  100  170
>
>1. Create a 95% confidence interval for the average calorie content in 200g of vanilla yoghurt.
>$$
> \begin{aligned}
>    \bar{x} &= 157.86\\
>    s &= 44.75\\
>    \\
>    \bar{x} \pm t^{*}_{df} \times \frac{s}{\sqrt{n}} &= 157.86 \pm 2.16 \times \frac{44.18}{\sqrt{14}}\\
>    &= 157.9 \pm 25.86\\
>    &= 132.0, 183.8\\
>  \end{aligned}
>$$
>    _i.e., we are 95% confident that the average calorie content ($\mu$) of all vanilla yoghurt is between 132.0 and 183.8 calories._
>
>
>2. A diet guide claims that there are 120 calories in a 200g serving of vanilla yoghurt. What does this evidence indicate?
>
>    _Since 120 calories is not contained in the interval (120 is below the lower bound of 132), the evidence indicates that the diet guide is incorrect and a 200g serve of the >        yoghurt has significantly more than 120 calories._

**GUIDED PRACTICE 7.12**

Using the information and results of Guided Practice 7.10 and Example 7.11, compute a 90% confidence interval for the average mercury content of croaker white fish (Pacific).

$$
  \begin{aligned}
    \bar{x} &= 0.287\\
    s &= 0.069\\
    df &= 15 - 1 = 14\\
    \\
    \bar{x} \pm t^{*}_{df} \times \frac{s}{\sqrt{n}} &= 0.287 \pm 1.76 \times \frac{0.069}{\sqrt{15}}\\
    &=  0.287 \pm 0.0314\\
    &= 0.256,\ 0.318\\
  \end{aligned}
$$
  _We are 90% confident that the average mercury content of croaker white fish (Pacific) is between 0.256 and 0.318 ppm._

**GUIDED PRACTICE 7.13**

The 90% confidence interval from Guided Practice 7.12 is 0.256 ppm to 0.318 ppm. Can we say that 90% of croaker white fish (Pacific) have mercury levels between 0.256 and 0.318 ppm?

  _No, a confidence interval only provides a range of plausible values for a population parameter, in this case the population mean. It does not describe what we might observe     for individual observations._
  
**Sample Question 2**

A random sample of 25 people followed a particular diet for 6 weeks. Their average weight loss was 5.6kg with standard deviation 0.8kg. Construct a 95% confidence interval for the average weight loss for this diet.
$$
  \begin{aligned}
    \bar{x} &= 5.6\\
    s &= 0.8\\
    \\
    \bar{x} \pm t^{*}_{df} \times \frac{s}{\sqrt{n}} &= 5.6 \pm 2.06 \times \frac{0.8}{\sqrt{25}}\\
    &=  5.6 \pm 0.3\\
    &= 5.3,\ 5.9\\
  \end{aligned}
$$
  _i.e., we are 95% confident that the average weight loss ($\mu$) for people following this diet is between 5.3 and 5.9 kg._
  
>## Inference for paired means
>  
>  $\bar{x}_d$ denotes the mean difference.
>  $s_d$ denotes the standard deviation differences.
>  
>  **Hypotheses**
>  $$
>    \begin{aligned}
>      H_0: \mu_d &= 0\\
>      H_a: \mu_d &< 0\ OR\\
>      \mu_d &< 0\ OR\\
>      \mu_d &\neq 0\
>    \end{aligned}
>  $$

>**Paired data**
>
>Which of the following situations would result in paired data:
>
>~~Comparing test results for one campus compared with another.~~
>
>$\checkmark$ Comparing individual students' Assignment 1 and 2 results.
>
>$\checkmark$ Investigating blood pressure change from a new drug.

>**Assumptions & Conditions**
>
>Differences must be independant of each other.
>
>Population differences ($N$) need to be a normal (distribution) model.
>
>A large sample size ($n$) will help to ensure normality.
>
>Random samples will ensure independence.

>**Confidence intervals for the mean difference ($\bar{x}_d$)**
>$$
>  \begin{aligned}
>    point\ estimate\ \pm t^{*}_{df} \times \frac{s_d}{\sqrt{n}}
>  \end{aligned}
>$$
>
>- $t^{*}_{df}$ will depend on the specified confidence level (i.e. 90%) and on the degrees of freedom ($n - 1$), and can be obtained from:
>   - t table
>   - qt() function in R

**t test statistic**

$$
  \begin{aligned}
    t = \frac{\bar{x}-\mu_0}{\frac{s_d}{\sqrt{n}}}
  \end{aligned}
$$

>**Hypotheses**
>
>A new drug designed to lower blood pressure is being trialed. A random sample of patients with high blood pressure took the drug for 6 weeks and their blood pressure was >recorded before taking the drug and at the end of the trial. Did their blood pressure decrease?
>
>Formulate the null and alternative hypotheses.
>$$
  \begin{aligned}
    H_0: \mu_d &= 0\\
    H_a: \mu_d &< 0\ (if\ d = after - before)\ OR\\
    H_a: \mu_d &> 0\ (if\ d = before - after)\\
  \end{aligned}
>$$

**Example**

A group of 9 randomly selected adults were given self defence lessons. Prior to the course, they were tested to determine their self-confidence. After the course they were given the same test. A high score on the test indicates a high degree of self-confidence. The scores and their differences (after – before) are given in the table below.

```{r, echo=FALSE}
  confidence <- matrix(c(6,10,8,6,5,4,3,8,5,8,12,9,6,7,5,4,9,5,2,2,1,0,2,1,1,1,0),ncol=3,byrow=FALSE)
  rownames(confidence) <- c("One","Two","Three","Four","Five","Six","Seven","Eight","Nine")
  colnames(confidence) <- c("Before","After","Difference")
  knitr::kable(confidence)
```

Note: $x_d=1.11$ and $s_d=0.78$

a. Do the results indicate that the course significantly increases the self-confidence of adults? Use $\alpha=0.01$.

$$
  \begin{aligned}
    H_0: \mu_d &= 0\\
    H_a: \mu_d &> 0\\
    \\
    t &= \frac{\bar{x}_d - \mu_0}{\frac{s}{\sqrt{n}}} = \frac{1.11 - 0}{\frac{0.78}{\sqrt{9}}}\\
    &= 4.27
  \end{aligned}
$$
P-value < 0.005

Since P-value $< \alpha$, reject $H_0$.

There is evidence that the self defence lessons significantly increase self-confidence in adults.

b.i. Calculate a 98% confidence for the true mean difference between the scores.

$$
  \begin{aligned}
    \bar{x} &= 1.11\\
    s &= 0.78\\
    \\
    \bar{x} \pm t^{*}_{df} \times \frac{s}{\sqrt{n}} &= 1.11 \pm 2.9 \times \frac{0.78}{\sqrt{9}}\\
    &=  1.11 \pm 0.754\\
    &= 0.36,\ 1.86\\
  \end{aligned}
$$

  _We are 98% sure than the true mean increase in self-confidence is between 0.36 and 1.86._

b.ii. Comment on your result in relation to your conclusion from a).
  _The confidence interval supports the hypothesis test result since 0 (no difference) is not contained in the interval and we rejected the null hypothesis that there was no     difference in self-confidence due to the self defence lessons._

>**Paired data activity**
>
>Complete the following questions with software or by hand.
>
>Twelve males aged between 13 and 16 were asked to self-report their height (in cm). Their heights were then measured. Is there sufficient evidence to support the claim that >there is a difference between self-reported and measured heights for males in this age group?

```{r, echo=FALSE}
  height <- matrix(c(173.0,180.0,160.0,178.0,180.0,152.0,165.0,162.0,137.0,160.0,168.0,183.0,172.5,177.5,164.8,173.5,178.6,153.9,163.8,170.2,141.2,188.5,165.1,179.8),ncol=2,byrow=FALSE)
  colnames(height) <- c("Reported","Mesured")
  knitr::kable(height)
```
>
>Note: $x_d=−2.617$ and $s_d=8.997$ (with d = reported − measured)
>
>a. Carry out a hypothesis test and use a 5% level of significance. 
>
>b. Report a 95% confidence interval to estimate the mean difference between reported heights and measured heights.
>
>c. How does your interval support your decision from the hypothesis test?
>